An optimization of bidirectional update language based on the idea of reversible computing \thanks{Supported by organization x.}

### 2019-08-29 paper/main.tex
update tex 2908
+ results and discussions

A restriction for $\pginline{case}$ needs to be introduced here. We know that there is one entering condition and one exit condition when evaluating $\putbxinline{case}$ as well as $\getbxinline{case}$. If a tupling $put$ and $get$ occurs, there will be 4 combinations of these conditions. This means that two entering conditions of $\putbxinline{case}$ and $\getbxinline{case}$ are not always simultaneously satified. The evaluated BXs are different in the $put$ and $get$ directions for combinations $((cond_{sv_1} \ s \ v) \ and \ (not (cond_{s_1} \ s)))$ and $((not (cond_{sv_1} \ s \ v) \ and \ (cond_{s_1} \ s))$, which are restricted in this paper. This does not happen for the others, which the construction of $\pginline{case}$ uses above. \\

@@

This section describes the experiments involving $minBIGUL$, $pg$, $cpg$, $kpg$ and $xpg$. The number of function calls, the evaluation time and the memory allocation are considered in each test.

@@

\begin{table}[]
    \centering
    \caption{Test cases}
    \label{tab:test-cases}
    \begin{tabular*}{\textwidth}{|c @{\extracolsep{\fill}}|l|l|l|c|c|c|c|}
        \hline
        \multirow{2}{*}{No} & \multicolumn{1}{c|}{\multirow{2}{*}{Name}} & \multicolumn{1}{c|}{\multirow{2}{*}{Associative}} & \multicolumn{1}{c|}{\multirow{2}{*}{Recursive}} & \multicolumn{2}{c|}{Input} & \multicolumn{2}{c|}{Output} \\ \cline{5-8} 
        & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{$s_0$} & \multicolumn{1}{c|}{$v_0$} & \multicolumn{1}{c|}{$s_r$} & \multicolumn{1}{c|}{$v_r$} \\ \hline
        1 & lassoc-comp-replace & left & no & 1 & 100 & 100 & 1 \\ \hline
        2 & rassoc-comp-replace & right & no & 1 & 100 & 100 & 1 \\ \hline
        3 & lassoc-comp-phead & left & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
        4 & rassoc-comp-phead & right & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
        5 & breverse & left & yes & [1..n] & [1..n] & [n..1] & [n..1] \\ \hline
    \end{tabular*}
\end{table}

@@

The last test case, $bReverse$, is a left-associative and recursive composition. It is defined in terms of $bFoldr$, a putback function for $foldr$ in BiGUL as well as minBiGUL. What we know about $foldr$, an important higher-order function on lists, is as follows:\\
    $\tab foldr :: (a \to b \to b) \to b \to [a] \to b\\
    \tab foldr \ f \ e \ [\ ] = e\\
    \tab foldr \ f \ e \ (x:xs) = f \ x \ (foldr \ f \ e \ xs)$

@@

The detailed explanation of $bFoldr$ are skipped. Like $phead$, the program equipvalent to the above $bFoldr$ is also provided in our OCaml system. If we put our eyes on the second branch of $bFoldr$, we can understand why it is recursive and left-associative composition. Note that the compose operator has a higher priority than the product operator. This leads to seemingly impossible to transform this function from left-associative style to right-associative style.

@@

\begin{table}[]
    \centering
    \caption{Comparison}
    \label{tab:comparison}
    \resizebox{\textwidth}{!}{%
        \begin{tabular}{|l|l|l|l|}
            \hline
            \multicolumn{1}{|c|}{Name} & \multicolumn{1}{c|}{Properties} & \multicolumn{1}{c|}{Evaluation time order} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Memory allocation order\\ (stack\_overflow point / 100000)\end{tabular}} \\ \hline
            lassoc-comp-replace & l, NR & xpg $\sim$ kpg $<$ cpg $<$ minbigul $<$ pg & minbigul $<$ cpg $<$ xpg $\sim$ kpg (95299) $<$ pg (ND) \\ \hline
            rassoc-comp-replace & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (95299) $<$ kpg (80638) \\ \hline
            lassoc-comp-phead & l, NR & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & minbigul $<$ xpg $\sim$ kpg (80638) $<$ cpg (ND) $<$ pg (ND) \\ \hline
            rassoc-comp-phead & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (43680) $<$ kpg (61665) \\ \hline
            breverse & l, R & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & xpg (ND) $<$ kpg (ND) $<$ cpg (ND) $<$ minbigul (ND) $<$ pg (ND) \\ \hline
            \multicolumn{4}{r}{} \\
            \multicolumn{4}{r}{stack\_overflow point (default: $>$ 100000): can not evaluate a program if the number of compositions is larger than this point} \\
            \multicolumn{4}{r}{best $< \ldots < \ldots <$ worst} \\
            \multicolumn{4}{r}{l/r: left/right associative} \\
            \multicolumn{4}{r}{NR/R: non-recursive/recursive} \\
            \multicolumn{4}{r}{ND: not determined}
        \end{tabular}%
    }
\end{table}

\begin{itemize}
    \item Number of calls
        \begin{itemize}
            \item No(cpg) = No(kpg) = No(xpg) = No(put) (\textcolor{blue}{come from function definitions})
            \item For right-associative \& non-recursive composition (rassoc-comp-replace, rassoc-comp-phead): No(get), No(put), No(pg) are linear where No(get) $<$ No(put) $<$ No(pg)
            \item For left-associative \& non-recursive composition (lassoc-comp-replace, lassoc-comp-phead): No(put) is still linear, No(get) is quadratic, No(pg) is under exponential distribution. No(put) $<$ No(get) $<$ No(pg)
            \item For left-associative \& recursive composition (breverse): we still have No(put) $<$ No(get) $<$ No(pg) and all three are non-linear. In case of breverse, No(put) is quadratic, No(get) is cubic and No(pg) is exponential.
        \end{itemize}
    \item Evalutation times (s)
        \begin{itemize}
            \item For left-associative compositions:\\
            - pg time is exponential (\textcolor{blue}{since the number of pg calls in these cases are exponential}) \\
            \item For left-associative \& non-recursive compositions:\\
            - kpg time are are approximate to xpg time, and they are the most efficient.\\
            - In case of lassoc-comp-replace, kpg time and xpg time are linear, cpg time and minbigul time are non-linear. (\textcolor{blue}{because the input size is constant, No(get) is quadratic for minbigul while the number of redundant evaluations in cpg is \underline{linear} and ... ?})\\
            - In case of lassoc-comp-phead, because of the data size is linear and the numbers of function calls are either linear or quadractic or exponential, then all times are non-linear.

            \item For right-associative \& non-recursive compositions:\\
            - minbigul time, pg time and xpg time are linear \\
            - In case of rassoc-comp-replace, all times are small enough to consider all them as linear \\
            - In case of rassoc-comp-phead, cpg time and kpg time are quadratic (\textcolor{blue}{because data size is linear, the number of redundant evaluations in cpg and kpg are linear})

            \item For left-associative \& recursive composition (breverse):\\
            - all times should be non-linear (\textcolor{blue}{due to there is no linear number of function calls})
        \end{itemize}
    \item Memory allocation (MBytes)
        \begin{itemize}
            \item Memory usage depends on the input size and the number of composition
            \item In general, minbigul uses less memory than the others for non-recursive compositions (all tests except breverse)
            \item pg only uses memory efficient when handling right-associative composition 
            \item cpg \& kpg are worst in memory allocation for right-associative
            \item For left-associative \& recursive composition (breverse), minbigul uses significantly more memory than kpg as well as xpg.
        \end{itemize}
\end{itemize}

@@

@@

>>>>

A restriction for $\pginline{case}$ needs to be introduced here. We know that there is one entering condition and one exit condition when evaluating $\putbxinline{case}$ as well as $\getbxinline{case}$. If a tupling $put$ and $get$ occurs, there will be 4 combinations of these conditions. This means that two entering conditions of $\putbxinline{case}$ and $\getbxinline{case}$ are not always simultaneously satified. The evaluated BXs are distinct in the $put$ and $get$ directions for combinations $((cond_{sv_1} \ s \ v) \ and \ (not (cond_{s_1} \ s)))$ and $((not (cond_{sv_1} \ s \ v) \ and \ (cond_{s_1} \ s))$, which are restricted in this paper. This does not happen for the others, which the construction of $\pginline{case}$ uses above. \\

@@

This section describes the experiments involving $minbigul$, $pg$, $cpg$, $kpg$ and $xpg$. The number of function calls, the evaluation time and the memory allocation are considered in each test.

@@

\begin{table}[]
    \centering
    \caption{Test cases}
    \label{tab:test-cases}
    \begin{tabular*}{\textwidth}{|c @{\extracolsep{\fill}}|l|l|l|c|c|c|c|}
        \hline
        \multirow{2}{*}{No} & \multicolumn{1}{c|}{\multirow{2}{*}{Name}} & \multicolumn{1}{c|}{\multirow{2}{*}{Associative}} & \multicolumn{1}{c|}{\multirow{2}{*}{Recursive}} & \multicolumn{2}{c|}{Input} & \multicolumn{2}{c|}{Output} \\ \cline{5-8} 
        & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{$s_0$} & \multicolumn{1}{c|}{$v_0$} & \multicolumn{1}{c|}{$s_r$} & \multicolumn{1}{c|}{$v_r$} \\ \hline
        1 & lassoc-comp-replace & left & no & 1 & 100 & 100 & 1 \\ \hline
        2 & rassoc-comp-replace & right & no & 1 & 100 & 100 & 1 \\ \hline
        3 & lassoc-comp-phead & left & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
        4 & rassoc-comp-phead & right & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
        5 & breverse & left/right & yes & [1..n] & [1..n] & [n..1] & [n..1] \\ \hline
    \end{tabular*}
\end{table}

@@

The last test case, $bReverse$, is a recursive composition. It is defined in terms of $bFoldr$, a putback function for $foldr$ in BiGUL as well as minbigul. What we know about $foldr$, an important higher-order function on lists, is as follows:\\
    $\tab foldr :: (a \to b \to b) \to b \to [a] \to b\\
    \tab foldr \ f \ e \ [\ ] = e\\
    \tab foldr \ f \ e \ (x:xs) = f \ x \ (foldr \ f \ e \ xs)$

@@

The detailed explanation of $bFoldr$ are skipped. Like $phead$, the program equipvalent to the above $bFoldr$ is also provided in our OCaml system. In the second branch of $bFoldr$, there is a composition of two BX programs where the first requires a recursive call. Then this composition can be classified as either left or right associative type. However it will look like a composition of more BX programs at some point if we slow down to evaluate it. Note that the compose operator has a higher priority than the product operator. This leads to seemingly impossible to transform the above composition from left-associative style to right-associative style.

@@

We will in turn consider the empirical results about the number of function calls, the evaluation time and the memory allocation.\\

Fig \ref{fig:count-calls} depicts the number of function calls including $put$, $get$, $pg$, $cpg$, $kpg$ and $xpg$ for each test case. From definitions of $put$, $cpg$, $kpg$ and $xpg$, we have No($cpg$) = No($kpg$) = No($xpg$) = No($put$) where No($f$) is the number of function call $f$. Here we consider that No($xpg$) in evaluating $\xpginline{bx_1 \circ bx_2}$ includes both the number of $kpg$s which are used in $\kpginline{bx_1}$ and the number of $xpg$s which are used in $\xpginline{bx_2}$.\\

For right-associative and non-recursive compositions (\textit{rassoc-comp-replace} and \textit{rassoc-comp-phead}), No($get$), No($put$) and No($pg$) are linear where No($get$) $<$ No($put$) $<$ No($pg$).\\

For left-associative and non-recursive compositions (\textit{lassoc-comp-replace} and \textit{lassoc-comp-phead}), No($put$) is still linear while No($get$) is quadratic and No($pg$) is under an exponential distribution. In this case, No($put$) $<$ No($get$) $<$ No($pg$).\\

For recursive composition (\textit{breverse}), the results still show that No($put$) $<$ No($get$) $<$ No($pg$) but all three are nonlinear. In case of \textit{breverse}, No($put$) is quadratic, No($get$) is cubic and No($pg$) is exponential.\\

@@

Fig \ref{fig:evaluation-time} illustrates the evaluation times in 5 test cases.\\

$pg$ times are always exponential for both left-associative compositions (\textit{lassoc-comp-replace}, \textit{lassoc-comp-phead}) and recursive compositions (\textit{breverse}) since the number of $pg$s in these cases are under exponential distributions.\\

For left-associative and non-recursive compositions, $kpg$ times are approximate to $xpg$ times, and they are the most efficient. In case of \textit{lassoc-comp-replace}, $kpg$ time and $xpg$ time are linear, $cpg$ time and $minbigul$ time are nonlinear. They are quite reasonable because the data sizes are constant, No($get$) is quadractic for minbigul while there is a linear redundant evaluation in $cpg$. In case of \textit{lassoc-comp-phead}, since the data sizes are linear and the numbers of function calls are either linear or quadractic or exponential, all times are nonlinear\\

For right-associative and non-recursive compositions, $minbigul$, $pg$ and $xpg$ times are linear. In case of \textit{rassoc-comp-replace}, all times are small enough to consider all them as linear. In case of \textit{rassoc-comp-phead}, both $cpg$ and $kpg$ times are quadratic because the data sizes are linear and the number of redundant evaluations are also linear.\\

For recursive compositions (\textit{breverse}), all times should be nonlinear since there is no linear number of function calls.\\

@@

Fig \ref{fig:memory-allocation} shows the memory usage which depends on the input size and number of compositions. In general, $minbigul$ uses less memory than the others for non-recursive compositions (all tests except \textit{breverse}). $pg$ only uses memory efficient when handling right-associative compositions. Also for evaluating these  compositions, both $cpg$ and $kpg$ are worst in memory allocation. For recursive compositions (\textit{breverse}), $minbigul$ uses significantly more memory than $kpg$ as well as $xpg$.

\begin{table}[]
    \centering
    \caption{Comparison}
    \label{tab:comparison}
    \resizebox{\textwidth}{!}{%
        \begin{tabular}{|l|l|l|l|}
            \hline
            \multicolumn{1}{|c|}{Name} & \multicolumn{1}{c|}{Properties} & \multicolumn{1}{c|}{Evaluation time order} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Memory allocation order\\ (stack\_overflow point / 100000)\end{tabular}} \\ \hline
            lassoc-comp-replace & l, NR & xpg $\sim$ kpg $<$ cpg $<$ minbigul $<$ pg & minbigul $<$ cpg $<$ xpg $\sim$ kpg (95299) $<$ pg (ND) \\ \hline
            rassoc-comp-replace & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (95299) $<$ kpg (80638) \\ \hline
            lassoc-comp-phead & l, NR & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & minbigul $<$ xpg $\sim$ kpg (80638) $<$ cpg (ND) $<$ pg (ND) \\ \hline
            rassoc-comp-phead & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (43680) $<$ kpg (61665) \\ \hline
            breverse & l/r, R & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & xpg (ND) $<$ kpg (ND) $<$ cpg (ND) $<$ minbigul (ND) $<$ pg (ND) \\ \hline
            \multicolumn{4}{r}{} \\
            \multicolumn{4}{r}{stack\_overflow point (default: $>$ 100000): can not evaluate a program if the number of compositions is larger than this point} \\
            \multicolumn{4}{r}{best $< \ldots < \ldots <$ worst} \\
            \multicolumn{4}{r}{l/r: left/right associative} \\
            \multicolumn{4}{r}{NR/R: non-recursive/recursive} \\
            \multicolumn{4}{r}{ND: not determined}
        \end{tabular}%
    }
\end{table}

Table \ref{tab:comparison} indicates a general comparison of the above results over the evaluation time and the memory allocation. This table provides some additional information about the point where a stack overflow error occurs in our OCaml system. We conducted tests while increasing the number of compositions from 10 to 100000. By default, the stack overflow point in the table is larger than 100000. In some cases, we limited the number of compositions because the corresponding evaluation times are too long. Then the above points are not determined.
